# Exploring mtcars dataset

### You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome).

--------------------------------------------------------------------------------------------------


## Data attributes
• (mpg)     Miles/(US) gallon.

• (cyl)     Number of cylinders.

• (disp)    Displacement (cu.in.).

• (hp)      Gross horsepower.

• (drat)    Rear axle ratio.

• (wt)      Weight (1000 lbs).

• (qsec)    1/4 mile time.

• (vs)      Engine (0 = V-shaped, 1 = straight).

• (am)      Transmission (0 = automatic, 1 = manual).

• (gear)    Number of forward gears.

• (carb)    Number of carburetors.


## Imports
```{r}
library("ggplot2")
library("reshape2")

```


# Adding the Dataset
```{r}

df = mtcars
print(df)

```

## The relation between Weight(wt) and Mile Per Gallon(mpg).
#### Conclusion: by increasing of car weight the distance travelled per gallon decreases.
```{r}

plt_1 = ggplot(df, aes(x=wt,y=mpg)) + geom_point()
print(
  plt_1 + 
    ggtitle("Increasing of car weight the distance travelled per gallon decreases") + 
    labs(y = "Mile Per Gallon Miles/(US) gallon", x = "Car Weight (1000 lbs)") 
  )

```


## The relation between Mile Per Gallon(mpg) and transmission(am) according to being automatic or manual .
#### 1 = Manual.
#### 0 = Automatic.
#### Conclusion: Manual cars travel more miles per gallon.
```{r}

plt_2 = ggplot(df,aes(x=wt,y=mpg,group=am)) +
        geom_point(aes(color=am)
        )

print(
  plt_2 +
    ggtitle("Manual cars travel more miles per gallon") +
    labs(y = "Mile Per Gallon Miles/(US) gallon", x = "Car Weight (1000 lbs)")  
      )


```

## Quantify the MPG difference between automatic and manual transmissions.
```{r}
# barplot 
# sum of Miles per gallon for manual cars
# sum of Miles per gallon for automatic cars

mpg_sum_auto <- sum(df[which(df$am==0), 1])
mpg_sum_manual <- sum(df[which(df$am==1), 1])
mpg_diff = mpg_sum_auto - mpg_sum_manual

ggplot(data=df, aes(x=am, y=mpg))+
  geom_bar(stat="identity", width=0.7, fill="steelblue") +
  ggtitle(paste("Sum of automatic MPG = ", mpg_sum_auto , " Miles\nSum of manual MPG = ", mpg_sum_manual , " Miles\nDifference = " , mpg_diff , " Miles")) +
  labs(y = "Total Miles Per Gallon Miles/(US) gallon", x = "Car Weight (1000 lbs)")  



```

## Relationships between multiple variables using corellation mx.
```{r}

cormat <- round(cor(df),2)
head(cormat)

```


### melt():
#### It enables us to reshape and elongate the data frames.
#### It organizes the data values in a long data frame format.
#### we need it in this elongated shape to plot the Heatmap.
```{r}

melted_cormat <- melt(cormat)
head(melted_cormat)

```


### We will use heatmap to analyze at least one dependent variable and three independent variables.
```{r}

ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value , label = value), color = "black", size = 4 ) + 
  
  geom_tile(color = "white") +
 scale_fill_gradient2(low = "deepskyblue1", high = "hotpink2", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Variable\nDependency\n(corelation)") +
  
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4)

```

#### Conclusion
Dependent variable: the variable will be dependent when the correlation is the largest.
Independent variable: the variable will be independent when the correlation is the smallest.

Dependent on mpg: Rear axle ratio, Engine (0 = V-shaped, 1 = straight), Transmission (0 = automatic, 1 = manual). 
Independent on mpg: weight.

```{r}
# for general dependency 

dependency = colMeans(cormat)
print(dependency)
# large number means dependent (correlated) on the rest of the features.
# small number means independent (uncorrelated) on the rest of the features.

```

## To find the best model fit to explain the relationship between the set of variables and miles per gallon (MPG) through a variable selection procedure called stepwise regression.
```{r}

# lock the randomization process
set.seed(42)

# choose random indices and set them as training/testing sets
n = nrow(mtcars)
index = sample(n, n*0.80, replace = FALSE)
train_set = mtcars[index,]
test_set = mtcars[-index,]

train_set
test_set

```

## Forward Selection
```{r}
# combination of regression
# mpg with < cyl , disp , hp , drat , wt , qsec , vs , am , gear , carb >

model_full = lm(mpg ~ ., data = train_set)
model_int = lm(mpg ~ -., data = train_set)

scopeformula = formula(model_full)
scopeformula

```
```{r}
fwd_sel = step(object=model_int, scope=scopeformula, direction="forward")
```
```{r}
summary(fwd_sel)
```

## Backward Selection
```{r}

model_full = lm(mpg ~ ., data = train_set)
scopeformula = formula(model_full)

back_sel = step(object=model_full, scope=scopeformula, direction="backward")

```
```{r}
summary(back_sel)
```

## Plotting for linear regression assumptions:
```{r}
plot(back_sel)
```













